# Lecture of DataFrames:
```{r Read data}
df1 <- read.table("audi.csv",header=T, sep=",")
df1$manufacturer <- "Audi"
df2 <- read.table("bmw.csv",header=T, sep=",")
df2$manufacturer <- "BMW"
df3 <- read.table("merc.csv",header=T, sep=",")
df3$manufacturer <- "Mercedes"
df4 <- read.table("vw.csv",header=T, sep=",")
df4$manufacturer <- "VW"

# Union by row:
df <- rbind(df1,df2,df3,df4)
dim(df)  # Size of data.frame
str(df) # Object class and description
names(df) # List of variable names

# Use birthday of 1 member of the group
set.seed(15042002)

# Random selection of x registers:
sam<-as.vector(sort(sample(1:nrow(df),5000)))
head(df)  # Take a look to the first rows/instances
df<-df[sam,]  # Subset of rows _ My sample
summary(df)

# Keep information in an .Rdata file:

save(list=c("df"),file="Cars.RData")
```
```{r}
summary (df)
```

#Let's see the classes of the variables
```{r}
class(df$model)
class(df$year)
class(df$price)
class(df$transmission)
class(df$mileage)
class(df$fuelType)
class(df$tax)
class(df$mpg)
class(df$engineSize)
class(df$manufacturer)
```

#If we want to load it, we only need to use `load`:

```{r Load data}
load("Cars.RData")
```

#We can transform variables into factors with `as.factor`:

```{r}
factormpg <- as.factor(df$mpg)
testVar <- as.factor(df$transmission)
summary(testVar)
```
```{r}
as.factor(df$mpg)
```

#Price Histogram

```{r}
hist(df$price, main="Price Histogram", xlab="Price", col="skyblue", border="black")
```
#Transmission Barplot

```{r}
barplot(table(df$transmission), main="Barplot of Transmission Types", xlab="Transmission Type", ylab="Frequency", col=c("skyblue", "lightpink"), border="black")

```
#Barplot of Fuel Types

```{r}
barplot(table(df$fuelType), main="Barplot of Fuel Types", xlab="Fuel Type", ylab="Frequency", col=c("skyblue", "lightpink", "lightgreen"), border="black")
```
#MPG Histogram

```{r}
hist(df$mpg, main="MPG Histogram", xlab="MPG", col="skyblue", border="black")
```
#Mileage Histogram

```{r}
hist(df$mileage, main="Mileage Histogram", xlab="Mileage", col="lightgreen", border="black")
```
#Model Barplot

```{r}
barplot(table(df$model), main="Barplot of df Models", xlab="Model", ylab="Frequency", col=rainbow(length(table(df$model))), border="black")
```

#Manufacturer Barplot

```{r}
barplot(table(df$manufacturer), main="Barplot of df Manufacturers", xlab="Manufacturers", ylab="Frequency", col=rainbow(length(table(df$manufacturer))), border="black")
```
#Year Barplot

```{r}
barplot(table(df$year), main="Barplot of Year", xlab="Year", ylab="Frequency", col=rainbow(length(table(df$year))), border="black")
```
#Tax Histogram

```{r}
hist(df$tax, main="Tax Histogram", xlab = "Tax", col = "lightyellow", border = "black")
```
#Engine Size Histogram

```{r}
hist(df$engineSize, main="Engine Size Histogram", xlab = "Engine Size", col = "lightcoral", border = "black")
```
```{r}
summary(df)
```

#We can transform variables into factors with `as.factor`:

#mpg:
```{r}
breaksmpg <- c(-Inf,  118.525,  235.95, 353.375, Inf)
labelsmpg <- c("very_inefficient", "inefficient", "efficient", "very_efficient")
new_mpg <- cut(df$mpg, breaks = breaksmpg, labels = labelsmpg)
summary(new_mpg)
plot(new_mpg, main="MPG Grouped Histogram",  col = rainbow(4))
```

#engineSize:
```{r}
breaksengineSize <- c(-Inf, 2.067, 4.13, Inf)
labelsengineSize <- c("small", "normal", "big")
new_engineSize <- cut(df$engineSize, breaks = breaksengineSize, labels = labelsengineSize)
summary(new_engineSize)
plot(new_engineSize, main="Engine Size Grouped Histogram",  col = rainbow(4))
```

#year:
```{r}
breaksyear <- c(-Inf, 2002.75, 2008.5, 2014.25, Inf)
labelsyear <- c("very_old", "old", "new", "very_new")
new_year <- cut(as.numeric(df$year), breaks = breaksyear, labels = labelsyear)
factoryear <- as.factor(new_year)
summary(new_year)
plot(new_year, main="Years Grouped Histogram",  col = rainbow(4))
```

#price:
```{r}
breaksprice <- c(-Inf, 35273.5, 68557, 101840.5, Inf)
labelsprice <- c("very_cheap", "cheap", "expensive", "very_expensive")
new_price <- cut(as.numeric(df$price), breaks = breaksprice, labels = labelsprice)
summary(new_price)
plot(new_price, main="Price Grouped Histogram" ,col = rainbow(4))
```

#mileage:
```{r}
breaksmileage <- c(-Inf, 64750.75, 129500.5, 194250.25, Inf)
labelsmileage <- c("very_used", "used", "shortly_used", "almost_new")
new_mileage <- cut(as.numeric(df$mileage), breaks = breaksmileage, labels = labelsmileage)
summary(new_mileage)
plot(new_mileage, main="Mileage Grouped Histogram",  col = rainbow(4))
```

#tax:
```{r}
breakstax <- c(-Inf, 145, 290, 435, Inf)
labelstax <- c("very_low", "low", "high", "very_high")
new_tax <- cut(as.numeric(df$tax), breaks = breakstax, labels = labelstax)
summary(new_tax)
plot(new_tax, main="Tax Grouped Histogram",  col = rainbow(4))
```

#Engine Size Histogram

```{r}
hist(df$engineSize, main="Engine Size Histogram", xlab = "Engine Size", col = "lightcoral", border = "black")
```
#Factorized Model Histogram

```{r}
factorModel <- as.factor(df$model)
summary(factorModel)
plot(factorModel, main="Factorized Model Histogram",  col = rainbow(24))
```
#Factorized Fuel Type Histogram

```{r}
factorFuelType <- as.factor(df$fuelType)
summary(factorFuelType)
#plot(factorFuelType, main="Factorized Fuel Type Histogram",  col = rainbow(24))
```
#Factorized Transmission Histogram

```{r}
factorTransmission <- as.factor(df$transmission)
summary(factorTransmission)
#plot(factorTransmission, main="Factorized Fuel Type Histogram",  col = rainbow(24))
```
#Factorized Manufactorer Histogram

```{r}
factorManufactorer <- as.factor(df$manufacturer)
summary(factorManufactorer)
#plot(factorTransmission, main="Factorized Fuel Type Histogram",  col = rainbow(24))
```

## Some useful functions

```{r, echo=FALSE}

# Some useful functions
calcQ <- function(x) {
  s.x <- summary(x)
  iqr<-s.x[5]-s.x[2]
  list(souti=s.x[2]-3*iqr, mouti=s.x[2]-1.5*iqr, min=s.x[1], q1=s.x[2], q2=s.x[3], 
       q3=s.x[5], max=s.x[6], mouts=s.x[5]+1.5*iqr, souts=s.x[5]+3*iqr ) }

countNA <- function(x) {
  mis_x <- NULL
  for (j in 1:ncol(x)) {mis_x[j] <- sum(is.na(x[,j])) }
  mis_x <- as.data.frame(mis_x)
  rownames(mis_x) <- names(x)
  mis_i <- rep(0,nrow(x))
  for (j in 1:ncol(x)) {mis_i <- mis_i + as.numeric(is.na(x[,j])) }
  list(mis_col=mis_x,mis_ind=mis_i) }

countX <- function(x,X) {
  n_x <- NULL
  for (j in 1:ncol(x)) {n_x[j] <- sum(x[,j]==X) }
  n_x <- as.data.frame(n_x)
  rownames(n_x) <- names(x)
  nx_i <- rep(0,nrow(x))
  for (j in 1:ncol(x)) {nx_i <- nx_i + as.numeric(x[,j]==X) }
  list(nx_col=n_x,nx_ind=nx_i) }

```

# Initialization of counts for missings, outliers and errors. All numerical variables have to be checked before and processing of outliers and errors
```{r}

dfVariables <- data.frame(price = df$price, engineSize = df$engineSize)


#######################################################
imis<-rep(0,nrow(df))  # rows - trips
jmis<-rep(0,ncol(df))  # columns - variables
######################################################
mis1<-countNA(df)
imis<-mis1$mis_ind
jmis<-mis1$mis_col$mis_x
barplot(as.numeric(imis), main = "Missings per Individual", col = rainbow(10))
barplot(as.numeric(mis1$mis_col$mis_x), main = "Missings per variable", col = rainbow(10))

#######################################################
iouts<-rep(0,nrow(df))  # rows - trips
jouts<-rep(0,ncol(df))  # columns - variables
######################################################

#######################################################
ierrs<-rep(0,nrow(df))  # rows - trips
jerrs<-rep(0,ncol(df))  # columns - variables
######################################################

#dfVariables <- subset(df, select = c(price, mileage, mpg, engineSize))


# Mileage
# It is a quantitive variable  Non-possible values will be recoded to NA
selMileage<-which(df$mileage <= 0)
ierrs[selMileage]<-ierrs[selMileage]+1
jerrs[5]<-length(selMileage)
selMileage                 #### sel contains the rownames of the individuals with "0" or less 
#                        as  value for mileage
df[selMileage,"mileage"]<-NA
selmissEngine <- selMileage

# EngineSize
# It is a quantitive variable  Non-possible values will be recoded to NA
selEngine<-which(df$engineSize <=0)
ierrs[selEngine]<-ierrs[selEngine]+1
jerrs[9]<-length(selEngine)
selEngine                 #### sel contains the rownames of the individuals with "0" 
#                        as  value for engineSize
# df[sel,"engineSize"]<-3    # non-possible values are replaced by NA, missing value symbol in R
df[selEngine,"engineSize"]<-NA
selmissEngine <- selEngine

# Price
# It is a quantitive variable  Non-possible values will be recoded to NA
selPrice<-which(df$price <=0)
ierrs[selPrice]<-ierrs[selPrice]+1
jerrs[3]<-length(selPrice)
selPrice                 #### sel contains the rownames of the individuals with "0" or less
#                        as  value for price
df[selPrice,"price"]<-NA
selmissPrice <- selPrice

# MPG
# It is a quantitive variable  Non-possible values will be recoded to NA
selMPG<-which(df$mpg < 0)
ierrs[selMPG]<-ierrs[selMPG]+1
jerrs[8]<-length(selMPG)
selMPG                 #### sel contains the rownames of the individuals with less than "0"
#                        as  value for MPG
# df[sel,"engineSize"]<-3    # non-possible values are replaced by NA, missing value symbol in R
df[selMPG,"mpg"]<-NA
selmissMPG <- selMPG

summary(df)
ierrs
numErrorsIndividuals<-which(ierrs>0)
numErrorsIndividuals
jerrs
barplot(jerrs, main="Errors per Variable", col = rainbow(10))
```

# Outlier detection and elimination
```{r}
boxplot(df$price, main="Boxplot of Price", ylab="Price")
var_outPrice<-calcQ(df$price)
summary(df$price)
abline(h=var_outPrice$souts,col="red")
abline(h=var_outPrice$souti,col="red")

lloutPrice<-which((df$price<var_outPrice$souti)|(df$price>var_outPrice$souts))
df$price[lloutPrice]<-NA
iouts[lloutPrice]<-iouts[lloutPrice]+1
jouts[3]<-length(lloutPrice)
boxplot(df$price, main="Clean Boxplot of Price", ylab="Price")

boxplot(df$engineSize, main="Boxplot of Engine Size", ylab="Engine Size")
var_outEngine<-calcQ(df$engineSize)
summary(df$engineSize)
abline(h=var_outEngine$souts,col="red")
abline(h=var_outEngine$souti,col="red")

lloutEngine<-which((df$engineSize<var_outEngine$souti)|(df$engineSize>var_outEngine$souts))
df$engineSize[lloutEngine]<-NA
iouts[lloutEngine]<-iouts[lloutEngine]+1
jouts[9]<-length(lloutEngine)
boxplot(df$engineSize, main="Clean Boxplot of Engine Size", ylab="Engine Size")

boxplot(df$mileage, main="Boxplot of Mileage", ylab="Mileage")
var_outMileage<-calcQ(df$mileage)
summary(df$mileage)
abline(h=var_outMileage$souts,col="red")
abline(h=var_outMileage$souti,col="red")

lloutMileage<-which((df$mileage<var_outMileage$souti)|(df$mileage>var_outMileage$souts))
df$mileage[lloutMileage]<-NA
iouts[lloutMileage]<-iouts[lloutMileage]+1
jouts[5]<-length(lloutMileage)
boxplot(df$mileage, main="Clean Boxplot of Mileage", ylab="Mileage")

boxplot(df$mpg, main="Boxplot of MPG", ylab="MPG")
var_outMPG<-calcQ(df$mpg)
summary(df$mpg)
abline(h=var_outMPG$souts,col="red")
abline(h=var_outMPG$souti,col="red")

lloutMPG<-which((df$mpg<var_outMPG$souti)|(df$mpg>var_outMPG$souts))
df$mpg[lloutMPG]<-NA
iouts[lloutMPG]<-iouts[lloutMPG]+1
jouts[5]<-length(lloutMPG)
boxplot(df$mpg, main="Clean Boxplot of MPG", ylab="MPG")

jouts
#df[llout,"engineSize"]<-NA
```

#Corr
```{r}
# Añadir nombres a los vectores jmis, jouts, y jerrs
variables <- c("model", "year", "price", "transmission", "mileage", "fuelType", "tax", "mpg", "engineSize", "manufacturer")
names(jmis) <- variables
names(jouts) <- variables
names(jerrs) <- variables

# 1. Describir estas variables y verificar qué otras variables tienen las asociaciones más altas.
variables_of_interest <- c("mileage", "mpg", "engineSize", "price")
desc_stats <- data.frame(
  Variable = variables_of_interest,
  Missing = imis[variables_of_interest],
  Outliers = iouts[variables_of_interest],
  Errors = ierrs[variables_of_interest]
)

# 2. Calcular la correlación con todas las otras variables y clasificar estas variables según la correlación.
correlations <- sapply(df[, variables_of_interest], function(x) cor(x, df$price, use="complete.obs"))
ranked_correlations <- sort(correlations, decreasing = TRUE)

# Mostrar gráfico de barras para las correlaciones
barplot(ranked_correlations, main="Correlaciones con Precio", ylab="Correlación", las=2, col="skyblue")

# 3. Calcular para cada grupo de individuos el promedio de valores faltantes/anómalos/errores para las variables de interés.
grouped_missing <- tapply(mis1$mis_ind, df$year, mean)
grouped_outliers <- tapply(iouts, df$year, mean) # assuming iouts is correctly calculated for outliers
grouped_errors <- tapply(ierrs, df$year, mean) # assuming ierrs is correctly calculated for errors

grouped_mean <- data.frame(
  Year = unique(df$year),
  Mean_Missing = grouped_missing,
  Mean_Outliers = grouped_outliers,
  Mean_Errors = grouped_errors
)

ranked_groups <- grouped_mean[order(-grouped_mean$Mean_Missing, -grouped_mean$Mean_Outliers, -grouped_mean$Mean_Errors), ]

# Mostrar gráfico de barras para el promedio de valores faltantes/anómalos/errores por año
barplot(ranked_groups$Mean_Missing, names.arg=ranked_groups$Year, main="Promedio de Valores Faltantes por Año", ylab="Promedio", las=2, col="lightgreen")
barplot(ranked_groups$Mean_Outliers, names.arg=ranked_groups$Year, main="Promedio de Anómalos por Año", ylab="Promedio", las=2, col="lightcoral")
barplot(ranked_groups$Mean_Errors, names.arg=ranked_groups$Year, main="Promedio de Errores por Año", ylab="Promedio", las=2, col="lightyellow")
```