The data for the project comes in a csv file:

```{r Read data}
bmw <- read.csv("bmw.csv")
```

Once we have our data read, we can save it as an RData file, which will be very easy to save and load:

```{r Save data}
save(bmw, file = "bmw.RData")
```

If we want to load it, we only need to use `load`:

```{r Load data}
load("bmw.RData")
```

We can transform variables into factors with `as.factor`:

```{r}
factormpg <- as.factor(bmw$mpg)
testVar <- as.factor(bmw$transmission)
summary(testVar)
```
```{r}
as.factor(bmw$mpg)
```

Price Histogram

```{r}
hist(bmw$price, main="Price Histogram", xlab="Price", col="skyblue", border="black")
```
Transmission Barplot

```{r}
barplot(table(bmw$transmission), main="Barplot of Transmission Types", xlab="Transmission Type", ylab="Frequency", col=c("skyblue", "lightpink"), border="black")

```
Barplot of Fuel Types

```{r}
barplot(table(bmw$fuelType), main="Barplot of Fuel Types", xlab="Fuel Type", ylab="Frequency", col=c("skyblue", "lightpink", "lightgreen"), border="black")
```
MPG Histogram

```{r}
hist(bmw$mpg, main="MPG Histogram", xlab="MPG", col="skyblue", border="black")
```
Mileage Histogram

```{r}
hist(bmw$mileage, main="Mileage Histogram", xlab="Mileage", col="lightgreen", border="black")
```
Model Barplot

```{r}
barplot(table(bmw$model), main="Barplot of BMW Models", xlab="Model", ylab="Frequency", col=rainbow(length(table(bmw$model))), border="black")
```
Year Barplot

```{r}
barplot(table(bmw$year), main="Barplot of Year", xlab="Year", ylab="Frequency", col=rainbow(length(table(bmw$year))), border="black")
```
Tax Histogram

```{r}
hist(bmw$tax, main="Tax Histogram", xlab = "Tax", col = "lightyellow", border = "black")
```
Engine Size Histogram

```{r}
hist(bmw$engineSize, main="Engine Size Histogram", xlab = "Engine Size", col = "lightcoral", border = "black")
```
```{r}
summary(bmw)
```

Engine Size Histogram

```{r}
hist(bmw$engineSize, main="Engine Size Histogram", xlab = "Engine Size", col = "lightcoral", border = "black")
```
#engineSize
factorModel <- as.factor(bmw$model)
summary(factorModel)
plot(factorModel,  col = rainbow(4)) 

